---
title: "Phase I"
output: html_notebook
---

# Load Libraries 
```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
```

# Load Data
```{r}
expert_df <- read.csv("expert_11.csv")
author_df <- read.csv("two_authors.csv")
df <- rbind(expert_df,author_df)
df
```
# Find items where there is at least 1 "No" for "Answerable"
## The resulting dataframe was used by the authors to discuss cases with disagreements and build the rulebook
```{r}
# Step 1: Identify trialName where there is at least one "No" for q3
trial_names_with_q3_no <- df %>%
  filter(Qid == 'q3', Answer == 'No') %>%
  select(trialName) %>%
  distinct()

# Step 2: Collect all rows in df with those trialNames regardless of "q3"
relevant_rows <- df %>%
  filter(trialName %in% trial_names_with_q3_no$trialName)

# Prepare for Step 3: Filter for only q3 and q5 answers and then pivot to widen the data
filtered_for_pivot <- relevant_rows %>%
  filter(Qid %in% c('q3', 'q5')) %>%
  select(Pid, trialName, Qid, Answer)

# Pivot to widen the data so q3 and q5 answers are in separate columns
final_df <- filtered_for_pivot %>%
  pivot_wider(names_from = Qid, values_from = Answer, names_prefix = "Answer_") %>%
  distinct()

# Rename the columns for clarity if necessary
colnames(final_df) <- c("Pid", "trialName", "q5", "q3")

# Display the resulting dataframe
arranged_final_df <- final_df %>% 
  arrange(trialName)
arranged_final_df # This dataframe was used by the authors to discuss cases with disagreements and build the rulebook
```
# Find items where there is at least 1 "No" or "More than one" for "Correctness"
## The resulting dataframe was used by the authors to discuss cases with disagreements and build the rulebook
```{r}
# Step 1: Find trialName where there's at least one "No" or one "More than one option could be considered correct" for q4
trialNames_with_q4_condition <- df %>%
  filter(Qid == 'q4', Answer %in% c('No', 'More than one option could be considered correct')) %>%
  distinct(trialName) %>%
  pull(trialName)

# Step 2: Collect all rows in df with those trialName regardless of "q4"
relevant_rows_df <- df %>%
  filter(trialName %in% trialNames_with_q4_condition, Qid %in% c('q4', 'q5')) 

# Step 3: Create a dataframe with Pid, trialName, q4, and q5
# Pivot wider to transform Qid values into separate columns for 'q4' and 'q5' answers
wide_df <- relevant_rows_df %>%
  select(Pid, trialName, Qid, Answer) %>%
  pivot_wider(names_from = Qid, values_from = Answer) %>%
  distinct() %>%
  select(Pid, trialName, q4 = `q4`, q5 = `q5`)

arranged_wide_df <- wide_df %>% 
  arrange(trialName)
arranged_wide_df # This dataframe was used by the authors to discuss cases with disagreements and build the rulebook
```

# Calculate the suggested Relevance rating for each item (part of the rulebook)
```{r}
# Make a copy of df
df_copy <- df

# Convert Answers in the copy to numeric where possible, errors will coerce to NA
df_copy$Answer <- as.numeric(as.character(df_copy$Answer))

# Filter for rows where Qid is 'q1', then group by 'trialName', and compute the average 'Answer' in the copy
average_q1_rating <- df_copy %>%
  filter(Qid == "q1") %>%
  group_by(trialName) %>%
  summarise(average_q1_rating = mean(Answer, na.rm = TRUE)) %>% 
  mutate(q1_floor = floor(average_q1_rating), q1_ceiling=ceiling(average_q1_rating))

average_q1_rating_for_phase_II <- average_q1_rating %>% 
  select(trialName, q1_floor, q1_ceiling)
average_q1_rating_for_phase_II
```

# Calculate the suggested Clarity rating for each item (part of the rulebook)
```{r}
# Make a copy of df
df_copy <- df

# Convert Answers in the copy to numeric where possible, errors will coerce to NA
df_copy$Answer <- as.numeric(as.character(df_copy$Answer))

# Filter for rows where Qid is 'q1', then group by 'trialName', and compute the average 'Answer' in the copy
average_q2_rating <- df_copy %>%
  filter(Qid == "q2") %>%
  group_by(trialName) %>%
  summarise(average_q2_rating = mean(Answer, na.rm = TRUE)) %>% 
  mutate(q2_floor = floor(average_q2_rating), q2_ceiling=ceiling(average_q2_rating))

average_q2_rating_for_phase_II <- average_q2_rating %>% 
  select(trialName, q2_floor, q2_ceiling)
average_q2_rating_for_phase_II
```

# Combine suggested ratings for Relevance and Clarity (part of the rulebook)
```{r}
ref_avg_rating <- average_q1_rating_for_phase_II %>% 
  left_join(average_q2_rating_for_phase_II, by = "trialName") 
ref_avg_rating$trialName <- sub("(.*)-[^-]*$", "\\1", ref_avg_rating$trialName)
ref_avg_rating
write.csv(ref_avg_rating, "rulebook_suggested_ratings.csv", row.names = FALSE)
```
